PHAT RAID, yo

### 2014-05-13 paper/paper.tex
Incomplete updates to evaluation

@@

\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{fit}
\usetikzlibrary{positioning}

@@

\newcommand{\chubby}[0]{\textsc{Chubby}}
\newcommand{\phat}[0]{\textsc{Phat}}
\newcommand{\phatraid}[0]{\textsc{PhatRaid}}
\newcommand{\raid}[1]{\textsc{RAID #1}}
\newcommand{\paxos}[0]{\textsc{Paxos}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\begin{figure}
  \begin{tikzpicture}
    [->,scale=.8,]
    \node [rectangle, draw, thick] (raidclient) at (0,2) {Raid Client};
    \node [rectangle, draw, thick] (client1) at (-4,1) {Client 1};
    \node [rectangle, draw, thick] (client2) at (0,0) {Client 2};
    \node [rectangle, draw, thick] (client3) at (4,1) {Client 3};
    \node [rectangle, draw, thick] (vr11) at (-7, -2) {Master};
    \node [rectangle, draw, thick] (vr12) at (-5, -2) {Replica};
    \node [rectangle, draw, thick] (vr13) at (-3, -2) {Replica};
    \node [rectangle, draw, thick] (vr21) at (-2, -4) {Master};
    \node [rectangle, draw, thick] (vr22) at (0, -4) {Replica};
    \node [rectangle, draw, thick] (vr23) at (2, -4) {Replica};
    \node [rectangle, draw, thick] (vr31) at (3, -2) {Master};
    \node [rectangle, draw, thick] (vr32) at (5, -2) {Replica};
    \node [rectangle, draw, thick] (vr33) at (7, -2) {Replica};

    \foreach \to in {client1, client2, client3}
    \draw [thick] (raidclient) -- (\to);
    \foreach \to in {vr11, vr12, vr13}
    \draw [thick] (client1) -- (\to);
    \foreach \to in {vr21, vr22, vr23}
    \draw [thick] (client2) -- (\to);
    \foreach \to in {vr31, vr32, vr33}
    \draw [thick] (client3) -- (\to);

@@

  \end{tikzpicture}
  \caption{Organizational Diagram}
  \label{fig:organization}
\end{figure}

@@

\subsection{Implenting RAID on \phat{}}

\subsection{Erlang and Behaviors}

@@

Listing \ref{lst:processing1} handles a message from an out-of-date master,
i.e., his commit number is older than the replica's commit number. The replica
first proposes a view change to the cluster. Afterwards, it changes its
\texttt{gen\_fsm} state to \lstinline!viewChange! and updates the
store\footnote{The store is called \lstinline!State! in the listings} with the
new view number and the new master.

@@

\section{Performance}

>>>>

\usepackage{caption}
\usepackage{subcaption}
% for subfigures

@@

\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{fit}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes.misc} % for cross shaped nodes

@@

\newcommand{\chubby}[0]{\textsc{Chubby}}
\newcommand{\phat}[0]{\textsc{Phat}}
\newcommand{\phatraid}[0]{\textsc{PhatRaid}}
\newcommand{\phatraidcf}[2]{\textsc{PhatRaid}(#1,#2)}
\newcommand{\raid}[1]{\textsc{RAID #1}}
\newcommand{\paxos}[0]{\textsc{Paxos}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Erlang Primer}

\phatraid{} is implemented in Erlang. Erlang is a pure, functional language with
extensive library support for distributed applications. Erlang has one unit of
concurrency, the process, and one unit of distribution, the node. Each Erlang
node consists of an arbitrary number of processes. Erlang nodes may live on the
same machine or on different machines connected by a network. Every Erlang node
is given a name by an Erlang name server. Erlang processes may communicate with
any Erlang node which they can name. Erlang processes interact with one another
by making RPC calls.

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    [<->,scale=.85,shorten <=2pt, shorten >=2pt,]
    \node [rectangle, draw, thick, inner sep=5pt] (raidclient) at (0,2) {Raid Client};
    \node [rectangle, draw, thick, inner sep=5pt] (client1) at (-4,1) {Client 1};
    \node [rectangle, draw, thick, inner sep=5pt] (client2) at (0,0) {Client 2};
    \node [rectangle, draw, thick, inner sep=5pt] (client3) at (4,1) {Client 3};
    \node [rectangle, draw, thick, inner sep=5pt] (vr11) at (-6, -2) {Replica};
    \node [rectangle, draw, thick, inner sep=5pt] (vr12) at (-4, -2) {Master};
    \node [rectangle, draw, thick, inner sep=5pt] (vr13) at (-2, -2) {Replica};
    \node [rectangle, draw, thick, inner sep=5pt] (vr21) at (-2, -4) {Replica};
    \node [rectangle, draw, thick, inner sep=5pt] (vr22) at (0, -4) {Master};
    \node [rectangle, draw, thick, inner sep=5pt] (vr23) at (2, -4) {Replica};
    \node [rectangle, draw, thick, inner sep=5pt] (vr31) at (2, -2) {Replica};
    \node [rectangle, draw, thick, inner sep=5pt] (vr32) at (4, -2) {Master};
    \node [rectangle, draw, thick, inner sep=5pt] (vr33) at (6, -2) {Replica};

    \foreach \to in {client1, client2, client3}
      \draw [thick] (raidclient) -- (\to);
    \draw [thick] (client1) -- (vr12);
    \draw [thick] (client2) -- (vr22);
    \draw [thick] (client3) -- (vr32);

    % \foreach \to in {vr11, vr13}
    %   \draw [thick] (vr12) -- (\to);
    % \foreach \to in {vr21, vr23}
    %   \draw [thick] (vr22) -- (\to);
    % \foreach \to in {vr31, vr33}
    %   \draw [thick] (vr32) -- (\to);

@@

  \end{tikzpicture}
  \caption{A \phatraidcf{3}{1} Group}
  \label{fig:organization}
\end{figure}

@@

\subsection{RAIDing \phat{}}

A \phatraidcf{$C$}{$f$} implementation is parameterized by two numbers: $C$, the
number of \phat{} clusters, and $f$, the \phat{} cluster failure threshold. A
single \phat{} cluster contains $2f + 1$ nodes. A \phatraid{} group contains $C$
clusters for a total of $C(2f+1)$ \phat{} nodes. A \phatraid{} group can be
concurrently accessed by an arbitrary number of raid-clients.

On a client node, the raid-client spawns $C$ client processes which each maintain
a connection to one of the $C$ \phat{} clusters. All communications between the
raid-client and a phat cluster go through the assigned client process.

\subsection{Erlang Behaviors}

@@

Listing \ref{lst:processing1} handles a message from an out-of-date master,
i.e., his commit number is older than the replica's commit number. The replica
first proposes a view change to the cluster. Afterwards, it changes its
\texttt{gen\_fsm} state to \lstinline!viewChange! and updates the
store\footnote{The store is called \lstinline!State! in the code listings} with
the new view number and the new master.

@@

\section{Evaluation}

\subsection{Resilience}

We distinguish between two types of failure: cessation of progress and loss of
data. In a VR cluster without persistent stores, data loss only occurs when all
$2f+1$ nodes in the cluster fail. In a VR cluster, progress only ceases if $f+1$
nodes are lost. In \phatraid{} the conditions for cessation of progress and loss
of data are complicated by the combination of RAID and VR.

\begin{figure}
  \centering
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[<->,scale=.5,shorten <=2pt, shorten >=2pt,]
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3, 3/1, 3/2, 3/3}
        \node [circle, draw, thick] (pr\x\y) at (\x,-\y) {};
      \foreach \x/\y in {3/1, 3/2, 3/3, 1/3, 2/3}
        \node [strike out, draw=red, thick] (pr\x\y) at (\x,-\y) {};
    \end{tikzpicture}
    \caption{Progress, No Data Loss}
    \label{fig:failures_allok}
  \end{subfigure}
~
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[<->,scale=.5,shorten <=2pt, shorten >=2pt,]
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3, 3/1, 3/2, 3/3}
        \node [circle, draw, thick] (pr\x\y) at (\x,-\y) {};
      \foreach \x/\y in {1/1, 1/2, 2/1, 2/2}
        \node [strike out, draw=red, thick] (pr\x\y) at (\x,-\y) {};
    \end{tikzpicture}
    \caption{No Progress, No Data Loss}
    \label{fig:failures_noprogress}
  \end{subfigure}
\\\vspace{1em}
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[scale=.5,shorten <=2pt, shorten >=2pt,]
      \node[rotate=25] (N) at (1.5,-1.5) {\textsc{Impossible!}};
    \end{tikzpicture}
    \caption{Progress, Data Loss}
    \label{fig:failures_max}
  \end{subfigure}
~
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[<->,scale=.5,shorten <=2pt, shorten >=2pt,]
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3, 3/1, 3/2, 3/3}
        \node [circle, draw, thick] (pr\x\y) at (\x,-\y) {};
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3}
        \node [strike out, draw=red, thick] (pr\x\y) at (\x,-\y) {};
    \end{tikzpicture}
    \caption{No Progress, Data Loss}
    \label{fig:failures_noprogress_nodata}
  \end{subfigure}

  \caption{Some Failed Node Distributions of \phatraidcf{3}{1}}
  \label{fig:failures}
\end{figure}

Four distributions of failed nodes in \phatraidcf{3}{1} are depicted in Figure
\ref{fig:failures}. In this section, if we claim ``data is lost'', we mean that
if no persistent store is used, data will be lost. Adding a persistent store
obviously prevents loss of committed data.

The top left corner depicts five of the nine nodes failing. This state is
\emph{not} a failure state. The group can still make progress and no data is
lost. The two surviving clusters can reach consensus on new values and can
reconstruct the failed cluster's data.

The top right corner depicts four of the nine nodes failing. This state is a
failure state. The group cannot make progress but no data can been
lost. Progress is blocked because more than one cluster cannot reach
consensus. Since no cluster was completely destroyed, all data still exists on
at least one node.

The bottom left corner is impossible because if data can be lost then we
certainly cannot make progress.

The bottom right corner depicts six of the nine nodes failing. This state is
also a failure state. The group cannot make progress and will lose
data. Progress is blocked because more than one cluster cannot reach
consensus. Additionally, two complete clusters have been lost so half of the
data cannot be recovered.

In general, progress ceases if more than one cluster cannot reach consensus and
data loss occurs if more than one cluster completely fails. The lower bound of
failed nodes for progress cessation in \phatraidcf{$C$}{$f$} is:
    $$ C\cdot(f+1) $$
and the lower bound of failed nodes for data loss is:
    $$ 2\cdot(2f+1) $$

In contrats, if all data has been propagated to all nodes, data is
only lost when all nodes die. If data has not been propagated to all nodes, then
loss occurs if all nodes containing the data die.

\subsection{Performance}
